Supplmental Information : Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-01-26 Ben_Manuscripts/structure_paper/supplemental.tex
new figures, better intro and more organization

  \vspace{1em}
  We parameterized monomers according to the following procedure:
  \begin{enumerate}
	\item \textit{Create monomer structure file with connectivity} : We
	drew atomistic structures using MarvinSketch
	17.13~\cite{chemaxon_marvinsketch_2017} with all hydrogen atoms drawn out
	explicitly. We optimized the 3D geometry of the structure using the 'Clean in
	3D' function of MarvinSketch.  We saved the structure as a .mol file, then
	converted it to .pdb format using Open Babel 2.4.1
	\cite{oboyle_open_2011,noauthor_open_nodate}. 
	\item \textit{Assign GAFF atomtypes using \texttt{antechamber}} : Using
	the .pdb structure file as input, we ran
	\texttt{antechamber}~\cite{wang_automatic_2006} using the AM1-BCC charge model.
	The net charge on the monomer is input as -1 since the sodium ion we keep it as
	a separate residue. We use \texttt{LEaP}~\cite{case_ambertools_nodate} and the
	output of \texttt{antechamber} to create Amber topology files. A detailed
	tutorial can be accessed elsewhere \cite{walker_antechamber_nodate}.
	\item \textit{Create gromacs topologies from Amber output} : The output
	of \texttt{LEaP} is a .inpcrd and a .prmtop file which are Amber topology
	files. Using acpype.py \cite{sousa_da_silva_acpype_2012}, we converted the
	\texttt{LEaP} output into GROMACS .gro and .top files. 
	\item \textit{Perform a simulated annealing procedure on the monomer} :
	We created a cubic box around the monomer using the GROMACS command \texttt{gmx
	editconf}. The monomer was centered in the box with edges of the
	box spaced at least 3 nm from the monomer on all sides. We ran an energy minimzation
	on the system with the steepest descent algorithm. Next we performed an NVT
	simulated annealing procedure. We linearly decreased the temperature of the
	system from 1000K to 50K over the course of 10 ns. We randomly chose a monomer
	configuration from the last 10 \% of the trajectory. 
	\item \textit{Reassign charges with \texttt{molcharge}}: With the monomer
	configuration taken from the annealed trajectory, we reassigned charges using
	\texttt{molcharge} with the am1bccsym method in order to ensure charges 
	are symmetric. This condition is not guaranteed with \texttt{antechamber}. 
	The charges in the GROMACS topology file (.top) were replaced with the 
	new charges calculated by \texttt{molcharge}. 
	\item \textit{Anneal again to get final structure} : We repeated the
	same simulated annealing procedure using the monomer topology with
	\texttt{molcharge} charges. A random monomer configuration was pulled from the
	last 10 \% of the trajectory and was used to build all assemblies reported
	(Figure~\ref{fig:monomer}).
  \end{enumerate}

@@

  \textbf{Calculation of pore-to-pore spacing statistics}

  We are interested in 5 pore-to-pore distances which should all be
  equal in a perfect hexagonal array (Figure~\ref{fig:p2p_diagram}). Only 4 of
  the 5 distances are independent. We can calculate a trajectory of spacing
  versus time for each of the 5 distances. We calculated the average pore-to-pore
  spacing and its uncertainty according to the following procedure:
   \begin{enumerate}
	\item We calculated the time when each of the pore-to-pore distances
	were equilibrated using \texttt{pymbar.timeseries.detectEquilibration()} \\
	\cite{chodera_simple_2016,shirts_statistically_2008}. We began calculations
	after the largest of the five values.
	\item We calculated how long it takes for the data in each of the 5
	trajectories to become uncorrelated using \texttt{pymbar.timeseries.integratedAutocorrelationTime()}
	\cite{chodera_simple_2016,shirts_statistically_2008}. 
	\item We broke the full equilibrated trajectory into blocks of length $\tau$, where
	$\tau$ is the maximum of the five autocorrelation times calculated. Each block contains
	five sub-trajectories of pore-to-pore spacings.   
	\item We generate statistics using the bootstrapping technique. For
	each bootstrap trial, we reconstruct an equilibrium trajectory by randomly
	sampling from the trajectory blocks. 
	\item The average pore-to-pore distance is the mean of all pore
	spacings among all bootstrap trials.
	\item To calculate the uncertainty in pore-to-pore distance, we calculate the average 
	pore spacing for each pore over all bootstrap trials. Using the 5 average 
	pore-to-pore distances, we calculate the spread with:
	\begin{align}
	s = \sqrt{\frac{1}{4} \sum_{i=1}^5 (x_i - \overline{x})^2} 
	\end{align}
	where $\overline{x}$ is the average pore-to-pore distance.
  \end{enumerate}

>>>>

  \vspace{1em}
  We parameterized monomers according to the following procedure:
  \begin{enumerate}
	\item \textit{Create monomer structure file with connectivity} : We
	drew atomistic structures using MarvinSketch
	17.13~\cite{chemaxon_marvinsketch_2017} with all hydrogen atoms drawn out
	explicitly. We optimized the 3D geometry of the structure using the 'Clean in
	3D' function of MarvinSketch.  We saved the structure as a .mol file, then
	converted it to .pdb format using Open Babel 2.4.1
	\cite{oboyle_open_2011,noauthor_open_nodate}. 
	\item \textit{Assign GAFF atomtypes using \texttt{antechamber}} : Using
	the .pdb structure file as input, we ran
	\texttt{antechamber}~\cite{wang_automatic_2006} using the AM1-BCC charge model.
	The net charge on the monomer is input as -1 since the sodium ion is kept as 
	a separate residue. We use \texttt{LEaP}~\cite{case_ambertools_nodate} and the
	output of \texttt{antechamber} to create Amber topology files. A detailed
	tutorial can be accessed elsewhere \cite{walker_antechamber_nodate}.
	\item \textit{Create GROMACS topologies from Amber output} : The output
	of \texttt{LEaP} is a .inpcrd and a .prmtop file which are Amber topology
	files. Using acpype.py \cite{sousa_da_silva_acpype_2012}, we converted the
	\texttt{LEaP} output into GROMACS .gro and .top files. 
	\item \textit{Perform a simulated annealing procedure on the monomer} :
	We created a cubic box around the monomer using the GROMACS command \texttt{gmx
	editconf}. The monomer was centered in the box with edges of the
	box spaced at least 3 nm from the monomer on all sides. We ran an energy minimzation
	on the system with the steepest descent algorithm. Next we performed an NVT
	simulated annealing procedure. We linearly decreased the temperature of the
	system from 1000 K to 50 K over the course of 10 ns. We randomly chose a monomer
	configuration from the last 10 \% of the trajectory. 
	\item \textit{Reassign charges with \texttt{molcharge}}: With the monomer
	configuration taken from the annealed trajectory, we reassigned charges using
	\texttt{molcharge} with the am1bccsym method in order to ensure charges 
	are symmetric. This condition is not guaranteed with \texttt{antechamber}. 
	The charges in the GROMACS topology file (.top) were replaced with the 
	new charges calculated by \texttt{molcharge}. 
	\item \textit{Anneal again to get final structure} : We repeated the
	same simulated annealing procedure using the monomer topology with
	\texttt{molcharge} charges. A random monomer configuration was pulled from the
	last 10 \% of the trajectory and was used to build all assemblies reported
	(Figure~\ref{fig:monomer}).
  \end{enumerate}

@@

  \noindent
  \begingroup
        \fontsize{14pt}{14pt}\selectfont
	\textbf{Calculation of pore-to-pore spacing statistics}
  \endgroup

  \vspace{1em} 
  We are interested in 5 pore-to-pore distances which should all be
  equal in a perfect hexagonal array (Figure~\ref{fig:p2p_diagram}). Only 4 of
  the 5 distances are independent. We can calculate a trajectory of spacing
  versus time for each of the 5 distances. We calculated the average pore-to-pore
  spacing and its uncertainty according to the following procedure:
   \begin{enumerate}
	\item We calculated the time when each of the pore-to-pore distances
	were equilibrated using \texttt{pymbar.timeseries.detectEquilibration()} \\
	\cite{chodera_simple_2016,shirts_statistically_2008}. We began calculations
	after the largest of the five values.
	\item We calculated how long it takes for the data in each of the 5
	trajectories to become uncorrelated using \texttt{pymbar.timeseries.integratedAutocorrelationTime()}
	\cite{chodera_simple_2016,shirts_statistically_2008}. 
	\item We broke the full equilibrated trajectory into blocks of length $\tau$, where
	$\tau$ is the maximum of the five autocorrelation times calculated. Each block contains
	five sub-trajectories of pore-to-pore spacings.   
	\item We generate statistics using the bootstrapping technique. For
	each bootstrap trial, we reconstruct an equilibrium trajectory by randomly
	sampling from the trajectory blocks. 
	\item The average pore-to-pore distance is the mean of all pore
	spacings among all bootstrap trials.
	\item To calculate the uncertainty in pore-to-pore distance, we calculate the average 
	pore spacing for each pore over all bootstrap trials. Using the 5 average 
	pore-to-pore distances, we calculate the spread with:
	\begin{align}
	s = \sqrt{\frac{1}{4} \sum_{i=1}^5 (x_i - \overline{x})^2} 
	\end{align}
	where $\overline{x}$ is the average pore-to-pore distance.
  \end{enumerate}
